Nonparametric Goodness-of-Fit Tests for Discrete Null Distributions

### 2010-11-15 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@27 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%General idea of goodness-of-fit tests (hypothesis test about underlying distribution).
%Given an observed sequence of data, the typical question of statistical inference is to determine
%the underlying process from which it came. 
Goodness-of-fit tests are used to assess whether data are consistent
with a hypothesized null distribution.  The $\chi^2$ test the best-known
parametric goodness-of-fit test, while the most popular nonparametric tests
are the classic test proposed by Kolmogorov and Smirnov followed closely by
several variants on an estimation procedure proposed by
Cram\'{e}r and von Mises (tests sometimes called Cram\'{e}r-von Mises-Smirnov
tests, or simply Cram\'{e}r-von Mises tests).

*** Citations in the paragraph above?  Also check the body of paper. ***

In their most basic forms, the nonparametric goodness-of-fit
tests are intended for continuous null distributions, but they have
also been adapted for use with discrete null distributions.  Unfortunately,
most modern statistical software have failed to incorporate these discrete
versions, leaving researchers with the $\chi^2$ test or in the unfortunate
position of applying a nonparametric test designed for continuous null
distributions and hoping for the best.
As we will see, the latter choice can be particularly dangerous
with small sample sizes. This paper presents a revision of R's
\code{ks.test()} function and a new \code{cvm.test()} function to
fill this void for researchers and practitioners in the R environment.

@@

Of all the methods for nonparametric goodness-of-fit tests, the most popular is the method devised by Kolmogorov and Smirnov.
It is the only such test built into the base of R. The idea behind it is fairly simple. Given the cumulative distribution
function $F_0(x)$ of the continuous null distribution, and the empirical distribution function $F_{data}(x)$ of the observed data, 
one constructs the statistic:
\begin{align}
D = \sup_x \left| F_0(x)- F_{data}(x) \right|
\end{align}
The distribution of $D$ under the null model does not depend on which null distribution is being used, making this a computationally
attractive method. For a standard treatment of the test and its performance relative to other algorithms see \cite{slakter65}. Two common
alternatives of the above test statistic exist. The absolute value is discarded and the rest is either left alone (the `greater' testing 
alternative) or the supremum is replaced with a negative infimum (the `lesser' hypothesis alternative). These can be helpful depending 
on the nature of the alternative hypotheses for which the test is desired to be powerful towards.

The extension of this result to non-continuous null distributions does not have such a clean solution. The formula of
the test statistic $D$ remains unchanged, however the distribution of the testing statistic is much more difficult. Unlike in the 
continuous case, it ultimately depends on which null model was chosen; this makes it impossible to simply read p-values directly 
off of a fixed table. It was known since at least the 1950's that using the tables for continuous distributions resulted in 
conservative p-values; it was not until \citet{Conover1972} that a method for computing exact p-values in this case was developed.

@@

The implementation of the discrete Kolmogorov-Smirnov function consists of two parts. First the particular test statistic needs to be calculated,
and then the p-value for that particular statistic must be computed.

Given that the test statistic is, theoretically, the same as in the
continuous case it would seem that the first part could be directly taken from the existing procedures. This is, unfortunately, often not the case.
Consider two non-decreasing functions $f$ and $g$, where the function $f$ is a step function with jumps on the set $\{x_1, \ldots x_N \}$ and $g$
is continuous. If we want to determine the supremum of the difference between these two functions notice that:

\begin{align}
\sup_x &\left| f(x)- g(x) \right|   \\
       &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
&=  \max_i \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg)
\end{align}

Computing the maximum over these $2N$ values (with $f$ equal to $F_{data}(x)$ and $g$ equal to $F_0(x)$ as defined above) is clearly the 
most efficient way to compute the Kolmogorov-Smirnov test statistic when given a continuous null distribution. When the function $g$ is not
continuous, notice that this formula no longer works since in general we cannot replace the limit of $g$ with its value at $x_i$. If it is known that $g$
is also a step function, we could replace the formula for some small $\epsilon$ by:
\begin{align}
\sup_x &\left| f(x)- g(x) \right| =   \\
        &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \}
\end{align}
Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function and knowing something
about its particular break-points. In the case of not knowing,
or having a $g$ which is in neither a step function nor a continuous function, the only method for computing the supremum is to take the numerical limit in (1). 
This clearly takes more computational time than simply cycling over $2N$ values.

Therefore, in order to implement the discrete Kolmogorov-Smirnov test statistic, we have forced the user to indicate the points of discontinuity
of the null distribution's cumulative distribution function. It is often the case that the test is used inside of a long simulation, 
and the added computational time would likely be prohibitive.

Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
continuous case. When the user requires exact p-values,
the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the R language, or in any other open source options, did not previously
exist and is included in the new function \pkg{ks.test}. The calculations are complex but straightforward; 
the full details are contained in the package source and in the original Conover paper.

@@

%General idea for continuous distributions. In many cases actually more powerful of the two. There again exists a distribution free 
%distribution for the testing statistic.
While the Kolmogorov-Smirnov test is the most well known of the non-parametric goodness of fit tests, there is another family of 
tests which has been shown to be more powerful to a large class of alternatives distributions. The original was developed jointly by
Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted by \cite{anderson1952}, and 
\cite{Watson1961}. The test statistics are, respectively, given as:
\begin{align}
W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
A^2 &= n \cdot \int_{-\infty}^{\infty} \frac{\left[F_{data}(x)- F_{0}(x) \right]^2}{F_0(x) -F_0(x)^2} dF_0(x) \\
U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x)
\end{align}
Where $F$ is either the cumulative distribution of the null model or the empirical cumulative distribution of the observed data.
As in the Kolmogorov-Smirnov test statistic, these all have distribution free null distributions in the continuous case.

The relative powers of these tests to different alternatives are studied in depth in \cite{stephens1974}. In general, the $W^2$
statistic is used unless there is a good reason not to. The $A^2$ statistic was developed by Anderson in order to be
generalized to the two-sample case, but was shown to perform very similarly to the original statistic in the one-sample
setting. Watson's $U^2$ statistic was developed for distributions which are cyclically distributed; that is they have some
order to them but no natural starting point. A common example would be the months of the year. His statistic will be same
if the data are cyclically reordered, since giving preference to one ordering would be unnatural.

@@

%Discuss the discrete version. How to calculate W2, U2, and A2. 
Generalizations of the Cram\'{e}r-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
the theoretical form of the test statistics are unchanged; although the discreteness allows for a slightly simpler representation.
The null distribution of the test statistics are again distribution dependent, unlike the continuous version. The methods do not
suggest finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
the original papers shows that this approximation is conservative and asymptotically equivalent to the true null distribution.

@@

Calculation of the three test statistics is done by straightforward matrix algebra as given in \cite{choulakian1994}. Determining 
the form of the asymptotic null distribution is also easy using the built-in eigenvalue decomposition functions. The only 
difficulty in the process involves actually calculating the percentiles for these weighted chi-squares.

%Show how to compute the distribution of a weighted sum of chi squared's.
The method used for calculating the distribution of a weighted sum of independent chi-squared variables is given in \cite{imhof1961}.
A general method for computing any quadratic form of normals is presented, which is easily adapted for our case since
each chi-squared variable has only one degree of freedom. The exact formula given for the distribution function of $Q$, the weighted
sum of chi squares, is:
\begin{align}
\mathbb{P}\{Q \geq x \} = \frac{1}{2} + \frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
\end{align}
For continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ which depend on the actual weights used.

There is no analytic solution to the integral, but integration can be carried out using numerical techniques. This seems
fine in most situations, but numerical issues do become a problem in the regime of large test statistics $x$. The function
$\theta(\cdot, x)$ is linear in $x$, and thus as the the test statistic grows the corresponding period of the integrand 
decreases. As the function acquires too many inflection points, the approximation becomes unstable. This is further
magnified by this occurring when $p$-values should be very small; thus tiny fluctuations which would be undetectable
elsewhere are quite prominent.

There is fortunately a simple conservative approximation which can get around this numerical problem. Given the
following inequality:
\begin{align}
\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \\
&= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
\end{align}
We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
of a higher degree of freedom. While for higher p-values it is better to use the original formulation, for small
p-values (we picked a cut-off of 0.001) this correction is useful and, given the small values, it should not
greatly affect the interpretation of the results.

Figure \ref{cvmissues} shows the non-monotonicity of the function as the test statistic grows, as well as how the
conservative approximation performs. While it is slightly conservative, it has the nice property of being strictly
monotone and not exhibiting other odd noisy behavior.

\begin{figure}
\begin{center}
\includegraphics[scale=0.4]{fig1.pdf}
\end{center}
\caption{Plot of p-value for given test statistics using numerical integration
(pink) versus a conservative chi-squared bound (blue). }
\label{cvmissues}
\end{figure}

@@

The revised \pkg{ks.test()} function supports one-sample tests for discrete
null distributions by allowing the second argument, \code{y}, to be
an empirical cumulative distribution function (an R function
with class \code{ecdf}) or an object of class \code{stepfun} specifying
a discrete distribution.  As in the original version of \code{ks.test()},
the presence of ties in the data (the first argument, \code{x}) generates a
warning unless \code{y} describes a discrete distribution.  When a discrete
distribution is specified, exact p-values are not available for 
two-sided alternative hypotheses, but the reported p-values will be
conservative.  For one-sided tests,
exact p-values are calculated using Conover's
method (when \code{exact = NULL} or \code{exact = TRUE})
if the sample size is less than or equal to 30; otherwise, asymptotic distributions
are used which are reliable in such cases (CITATION? Is this correct?).
When \code{exact = FALSE} the asymptotic distribution is used which is known
to be imprecise but conservative, even for small samples (CITATION).

@@

The chi-squared test is a popular alternative to the methods presented here for doing goodness-of-fit tests for discrete data.
It differs by using no information about the geometry of the distribution, instead relying only the difference between the 
number of data points observed at each value and the expected number of data points that should be observed at each point. 
This generally has the effect of having a greater power towards alternatives with sharply different densities on a few points,
at the expense of losing power for alternatives with slight deviations over all of the points. Also, the non-parametric tests
have an advantage of being useful when the sample size is lower compared to the support of the distribution. The chi-squared
test is generally not used unless there are at least five data points in each element of the support. On the other hand,
chi-squared can be used in situations when there is not a numerical interpretation of the observation space 
(e.g. a set of ethnicities).
For a more complete comparison of the two tests' power see \cite{slakter65}.

The computational capabilities of modern computers provide an alternative to using a complex formula to
calculate the p-value for a test statistic. By drawing random samples from the null distribution, in many
cases the p-values for a given statistic can be calculated accurately in a relatively short time span.
This can be quite useful in some cases, but using the exact formulas from our methodology has several
distinct advantages. First of all, if the function is called many times (say, within another simulation
study), the computational benefits of using a formula can quickly become substantial. There is also a greater need
for supervision in a simulation study, with the exact number of runs needed to reach convergence (or other 
similar threshold). Additionally, while
simulations are generally a well received accepted method, it is often important in applied data analysis
to be sure that the calculated $p$-values are not the result of a numerical oddity in a particular
run of a simulation.

@@

In the continuous setting, both the Kolmogorov-Smirnov and the Cram\'{e}r-von Mises tests have two sample analogues. Here data are
observed from two processes, and the hypothesis tested is whether they came from the same (but unspecified) distribution. There
does not exist an analogous theory for discrete distributions. This comes from the fact that the discrete null distributions
of the test statistics depend on the exact null distribution; therefore the two sample case would surely have to depend
on the exact distributions used as well, which are generally not even stated in the two-sample case.

>>>>

%General idea of goodness-of-fit tests (hypothesis test about underlying distribution).
%Given an observed sequence of data, the typical question of statistical inference is to determine
%the underlying process from which it came. 
Goodness-of-fit tests are used to assess whether data are consistent
with a hypothesized null distribution.  The $\chi^2$ test is the best-known
parametric goodness-of-fit test, while the most popular nonparametric tests
are the classic test proposed by Kolmogorov and Smirnov followed closely by
several variants on an estimation procedure proposed by
Cram\'{e}r and von Mises (tests sometimes called Cram\'{e}r-von Mises-Smirnov
tests, or simply Cram\'{e}r-von Mises tests).

*** why "estimation procedure" above?
*** Citations in the paragraph above?  Also check the body of paper. ***

In their most basic forms, these nonparametric goodness-of-fit
tests are intended for continuous null distributions, but they have
also been adapted for use with discrete null distributions.  Unfortunately,
most modern statistical software have failed to incorporate these discrete
versions, leaving researchers with the $\chi^2$ test or in the unfortunate
position of applying a nonparametric test designed for a continuous null
distribution and hoping for the best.
As we will see, this latter course of action can be particularly dangerous
with small sample sizes. This paper presents a revision of R's
\code{ks.test()} function and a new \code{cvm.test()} function to
fill this void for researchers and practitioners in the R environment.

*** Later, make sure we point out how using one of the ``old'' tests and
hoping for the best is a problem. ***

@@

The most popular nonparametric goodness-of-fit test is the
Kolmogorov-Smirnov test. The idea is fairly simple:
Given the cumulative distribution
function $F_0(x)$ of the continuous null distribution
and the empirical distribution function $F_{data}(x)$ of the observed data,  the test statistic is given by
\begin{align}
D = \sup_x \left| F_0(x)- F_{data}(x) \right|.
\end{align}
The distribution of $D$ does not depend on the hypothesized
distribution, making this a computationally
attractive method. \cite{slakter65} offers a standard presentation
of the test and its
performance relative to other algorithms. 
The test statistic (1) is easily adapted for one-sided tests.
For these, the absolute value is discarded and the tests are based
on either the supremum of the remaining difference (the `greater' testing
alternative) or by replacing the supremum with a negative infimum
(the `lesser' hypothesis alternative).  Tabulated p-values have been
available for these tests since (*** year ****).

**** check this above, both do away with the absolute value, I think? ***

The extension of the Kolmogorov-Smirnov test to non-continuous 
null distributions does not have such a clean solution. The formula of
the test statistic $D$ remains unchanged, but the distribution
of the test statistic is much more difficult to obtain. Unlike the 
continuous case, the null distribution depends on the null model.
It has been known since at least the 1950's that using the tables associated
with continuous null distributions results in conservative p-values
(**** citation?) when the null distribution is discontinuous 
(*** or should this be discrete? ***).  In the early 1970's, 
\citet{Conover1972} developed the method implemented here
for computing exact p-values
in the case of discrete null distributions.

@@

The implementation of the discrete Kolmogorov-Smirnov test is considered
in two parts. First, the particular test statistic needs to be calculated
(corresponding to the desired one-sided or two-sided test).
Then, the p-value for that particular test statistic may be computed.

The test statistic is, in terms of the theory, the same as in the
continuous case; it would seem that no additional work would be required,
but unfortunately this is not the case.
Consider two non-decreasing functions $f$ and $g$, where the function $f$ is a step function with jumps on the set $\{x_1, \ldots x_N \}$ and $g$
is continuous (the classical Kolmogorov-Smirnov situation).
In order to determine the supremum of the difference between
these two functions, notice that
\begin{align}
\sup_x &\left| f(x)- g(x) \right|  \nonumber  \\
       &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
&=  \max_i \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg).
\end{align}
Computing the maximum over these $2N$ (*** why 2N? ***)
values (with $f$ equal to
$F_{data}(x)$ and $g$ equal to $F_0(x)$ as defined above) is clearly the 
most efficient way to compute the Kolmogorov-Smirnov test statistic for
a continuous null distribution. When the function $g$ is not
continuous, however, equality (3) does not hold in general because 
we cannot replace $\lim_{x\rightarrow x_e} g(x)$ with the value $g(x_i)$.

If it is known that $g$ is a step function, it follows that
for some small $\epsilon$,
\begin{align}
\sup_x &\left| f(x)- g(x) \right| =  \nonumber \\
        &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \}
\end{align}
where the discontinuities in $g$ are at least a distance $\epsilon$ apart. 
This, however, requires knowledge that $g$ is a step function as well as of
the nature of its support (or break-points).
%
%*** Hmm... do we care about time? ***  When the break-points of $g$ are
%unknown, or when $g$ is neither a step function nor a continuous distribution
%function, computing the supremum requires taking the numerical limit in (1***?). 
%This is computational involved and risks problems with numerical tolerances.
%
We implement the Kolmogorov-Smirnov test statistic for discrete null
distributions by forcing the user to completely specify the null distribution.

%**** Hmmm... It is often the case that the test is used inside of a
%long simulation, 
%and the added computational time would likely be prohibitive.

Having obtained the test statistic, the p-value must then be calculated. 
For larger sample sizes, the null distribution coincides 
(***should this be
stated precisely as an asymptotic result???)
with the standard null distribution (CITATION???). 
When an exact p-value is required for
smaller sample sizes, the methodology in \citet{Conover1972} is used. 
Full details of the calculations are contained in source code
of our revised function \code{ks.test()} and in Conover's original paper.

** code hasn't previously existed?  I deleted this stuff, not sure if
it fits here.  Is it obvious (since we bothered to do this in the first
place)?  Or move it?

%Code for this procedure in the R language, or in any other open source options, did not previously
%exist and is included in the new function \pkg{ks.test}.

@@

%General idea for continuous distributions. In many cases actually more powerful of the two. There again exists a distribution free 
%distribution for the testing statistic.

While the Kolmogorov-Smirnov test may be the most popular of
the nonparametric goodness-of-fit tests, there is another family of 
tests which has been shown to be more powerful against a large class
of alternatives hypotheses. 
The original test was developed jointly (*** or was it simultaneously? ***) by
Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928} and further adapted by \cite{anderson1952}, and  \cite{Watson1961}. 
The original test statistic, $W^2$, Anderson's $A^2$, and Watson's
$U^2$ are:
\begin{align}
W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
A^2 &= n \cdot \int_{-\infty}^{\infty} \frac{\left[F_{data}(x)- F_{0}(x) \right]^2}{F_0(x) -F_0(x)^2} dF_0(x) \\
U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x)
\end{align}
As with the original Kolmogorov-Smirnov test statistic, these all have 
null distributions of their test statistics which are independent of the
specified null models.

The relative powers of these tests to different alternatives are
studied in depth in \cite{stephens1974}. In general, the $W^2$
statistic is recommended unless there is a compelling reason to
consider one of the others. The $A^2$ statistic was developed by
Anderson in the process of generalizing the test for the two-sample case,
but was shown to perform very similarly to the original statistic
in the one-sample setting (*** so why, then would we ever use it??? ***).
Watson's $U^2$ statistic was developed for distributions
which are cyclic (with an ordering to the support but
no natural starting point  *** how best to describe this? ***). 
For example, a distribution on the months of the year could be considered
cyclic. Watson's statistic is invariant to cyclical reordering (giving preference to one ordering would be unnatural).  **** is this last point necessary? ***

*** Is a second paragraph on Stephen's results really necessary?  Can we
pick and choose and get away with one paragraph? ***

@@

%Discuss the discrete version. How to calculate W2, U2, and A2.

*** I edited, below, make sure you check it. ****

Generalizations of the Cram\'{e}r-von Mises tests to discrete
distributions were developed in \cite{choulakian1994}. As with the
Kolmogorov-Smirnov test, the forms of the test statistics are unchanged, although the discreteness allows for slightly simpler representations.
The null distribution of the test statistics are again hypothesis-dependent. The methods do not offer finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
chi-squared variables (with the weights depending on the particular
null distribution). (*** is this true for all 3? ***)
We implement this asymptotic distribution here;
the original papers (*** really, all of them? ***)
show that this approximation is conservative 
and asymptotically equivalent to the true null distribution.

*** What does it mean for a distribution to be equal to a weighted sum of
independent chi-squared variables?  I think this needs to be stated more
precisely, above.  A formal definition of $Q$ is needed someplace, too,
as I'll probably note below. ***

@@

Calculation of the three test statistics is done using the
matrix algebra given by \cite{choulakian1994}. The form of the
asymptotic null distribution is obtained using R's
eigenvalue decomposition functions.

The only notable difficulty in the process involves calculating the percentiles
of the weighted sum of chi-squares,
\begin{align}
Q = \sum_{i=1}^{p} \lambda_i \chi^2_{i,1df}
\end{align}
\cite{imhof1961} provides a method for obtaining the distribution of
a weighted sum of independent chi-squared variables.  It is
easily adapted for our case because
the chi-squared variables have only one degree of freedom.
The exact formula given for the distribution function of $Q$
is given by
\begin{align}
\mathbb{P}\{Q \geq x \} = \frac{1}{2} + 
\frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
\end{align}
for continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ depending on the actual weights used.

There is no analytic solution to the integral, so the integration is
accomplished numerically. This seems fine in most situations we considered,
but numerical issues appear in the regime of large test statistics $x$
(or, equivalently, small p-values).
The function $\theta(\cdot, x)$ is linear in $x$; as the test statistic 
grows the corresponding periodicity of the integrand decreases and
the approximation becomes unstable. 
%This is further magnified in situations where the 
%$p$-values should be very small, where tiny fluctuations (undetectable
%elsewhere) become quite prominent. 
We resolve this problem by using a simple conservative approximation to
avoid the numerical instability. Consider the
following inequality:
\begin{align}
\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \\
&= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
\end{align}
The values for the weighted sum can be bounded using a simple transformation
and a chi-squared distribution of a higher degree of freedom. 
The original formulation is preferable for most p-values, while
this approximation is useful for smaller p-values (smaller than 0.001,
based on our observations of the numerical instablity of the original
formulaton); it should not greatly affect the interpretation of the results.

*** How is the cutoff established in the code? ***

Figure \ref{cvmissues} shows the non-monotonicity of the function 
(*** which function, reference it? ***) as the test statistic grows, 
compared to the conservative approximation.

*** stuff deleted, redundant ***.

\begin{figure}
\begin{center}
\includegraphics[scale=0.4]{fig1.pdf}
\end{center}
\caption{Plot of p-value for given test statistics using numerical integration
(pink) versus a conservative chi-squared bound (dashed blue). }
\label{cvmissues}
\end{figure}

@@

The revised \code{ks.test()} function supports one-sample tests for discrete
null distributions by allowing the second argument, \code{y}, to be
an empirical cumulative distribution function (an R function
with class \code{ecdf}) or an object of class \code{stepfun} specifying
a discrete distribution.  As in the original version of \code{ks.test()},
the presence of ties in the data (the first argument, \code{x}) generates a
warning unless \code{y} describes a discrete distribution.  When a discrete
distribution is specified, exact p-values are not available for 
two-sided alternative hypotheses, but the reported p-values will be
conservative.  For one-sided tests,
exact p-values are calculated using Conover's
method (when \code{exact = NULL} or \code{exact = TRUE})
if the sample size is less than or equal to 30; otherwise, asymptotic distributions
are used which are reliable in such cases (CITATION? Is this correct?).
When \code{exact = FALSE} the asymptotic distribution is used which is known
to be imprecise but conservative, even for small samples (CITATION).

@@

*** I commented out a bunch of stuff.  I think the discussion can be short
and sweet. ***

%The chi-squared test is a popular alternative to the methods presented here for doing goodness-of-fit tests for discrete data.
%It differs by using no information about the geometry of the distribution, instead relying only the difference between the 
%number of data points observed at each value and the expected number of data points that should be observed at each point. 
%This generally has the effect of having a greater power towards alternatives with sharply different densities on a few points,
%at the expense of losing power for alternatives with slight deviations over all of the points. Also, the non-parametric tests
%have an advantage of being useful when the sample size is lower compared to the support of the distribution. The chi-squared
%test is generally not used unless there are at least five data points in each element of the support. On the other hand,
%chi-squared can be used in situations when there is not a numerical interpretation of the observation space 
%(e.g. a set of ethnicities).
%For a more complete comparison of the two tests' power see \cite{slakter65}.

%The computational capabilities of modern computers provide an alternative to using a complex formula to
%calculate the p-value for a test statistic. By drawing random samples from the null distribution, in many
%cases the p-values for a given statistic can be calculated accurately in a relatively short time span.
%This can be quite useful in some cases, but using the exact formulas from our methodology has several
%distinct advantages. First of all, if the function is called many times (say, within another simulation
%study), the computational benefits of using a formula can quickly become substantial. There is also a greater need
%for supervision in a simulation study, with the exact number of runs needed to reach convergence (or other 
%similar threshold). Additionally, while
%simulations are generally a well received accepted method, it is often important in applied data analysis
%to be sure that the calculated $p$-values are not the result of a numerical oddity in a particular
%run of a simulation.

*** Hmm... if you claim the following, you want to make sure it is true.
Did you read it somewhere? ***

@@

In the continuous setting, both of 
the Kolmogorov-Smirnov and the Cram\'{e}r-von Mises tests have two-sample analogues. Here data are
observed from two processes, and the hypothesis tested is 
whether they came from the same (but unspecified) distribution. There
does not exist an analogous theory for discrete distributions. 
This comes from the fact that the discrete null distributions
of the test statistics depend on the exact null distribution; 
therefore the two-sample case would surely have to depend
on the exact distributions used as well (*** is this true? ***), 
which are generally not even stated in the two-sample case.
